---
title: "Exponential Distribution and the Central Limit Theorem"
author: "jameel"
date: "2024-05-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this report, we investigate the properties of the exponential distribution and compare it with the Central Limit Theorem (CLT). Specifically, we will explore the distribution of the averages of 40 exponentials through simulation.


## Simulation and Analysis


```{r  echo=FALSE}
# Set the parameters
lambda <- 0.2
n <- 40
num_simulations <- 1000

# Simulate the exponential distribution
set.seed(123)  # Setting seed for reproducibility
simulations <- replicate(num_simulations, rexp(n, lambda))

# Calculate the sample means
sample_means <- colMeans(simulations)

# Theoretical mean and variance
theoretical_mean <- 1 / lambda
theoretical_variance <- (1 / lambda)^2 / n

# Sample mean and variance
sample_mean <- mean(sample_means)
sample_variance <- var(sample_means)

# Display the results
cat("Sample Mean:", sample_mean, "\n")
cat("Theoretical Mean:", theoretical_mean, "\n")
cat("Sample Variance:", sample_variance, "\n")
cat("Theoretical Variance:", theoretical_variance, "\n")

```

## Sample Mean
The sample mean of the distribution of 40 exponentials is r sample_mean, while the theoretical mean is r theoretical_mean. These values are very close, illustrating the Law of Large Numbers.

## Sample Variance 

The sample variance is r sample_variance, which is also close to the theoretical variance of r theoretical_variance.

## Distribution of Sample Means
The histogram below shows the distribution of the sample means. We also overlay the theoretical normal distribution to illustrate the Central Limit Theorem.

```{r  echo=FALSE}
hist(sample_means, probability = TRUE, 
     main = "Distribution of Sample Means of 40 Exponentials",
     xlab = "Sample Mean", 
     col = "lightblue", 
     border = "black")
lines(density(sample_means), col = "red", lwd = 2)
x <- seq(min(sample_means), max(sample_means), length.out = 100)
y <- dnorm(x, mean = theoretical_mean, sd = sqrt(theoretical_variance))
lines(x, y, col = "blue", lwd = 2, lty = 2)


```


## Normality Check 
The QQ-plot below shows that the distribution of the sample means is approximately normal.

`
```{r  echo=FALSE}
qqnorm(sample_means, main = "QQ-Plot of Sample Means")
qqline(sample_means, col = "red")

```

## Conclusion
The sample mean and variance of the distribution of 40 exponentials closely match the theoretical values. The distribution of the sample means is approximately normal, as expected from the Central Limit Theorem.

## Apendix
```{r  echo=FALSE}
# Display the code used in the simulation
lambda <- 0.2
n <- 40
num_simulations <- 1000
set.seed(123)
simulations <- replicate(num_simulations, rexp(n, lambda))
sample_means <- colMeans(simulations)
theoretical_mean <- 1 / lambda
theoretical_variance <- (1 / lambda)^2 / n
sample_mean <- mean(sample_means)
sample_variance <- var(sample_means)
hist(sample_means, probability = TRUE, 
     main = "Distribution of Sample Means of 40 Exponentials",
     xlab = "Sample Mean", 
     col = "lightblue", 
     border = "black")
lines(density(sample_means), col = "red", lwd = 2)
x <- seq(min(sample_means), max(sample_means), length.out = 100)
y <- dnorm(x, mean = theoretical_mean, sd = sqrt(theoretical_variance))
lines(x, y, col = "blue", lwd = 2, lty = 2)
qqnorm(sample_means, main = "QQ-Plot of Sample Means")
qqline(sample_means, col = "red")


```



