---
title: "Analysis of ToothGrowth Data"
author: "jameel"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this report, we analyze the ToothGrowth dataset from the R datasets package. We will perform basic exploratory data analyses, provide a summary of the data, and use confidence intervals and hypothesis tests to compare tooth growth by supplement type (supp) and dose.


## Exploratory Data Analysis


```{r, echo=FALSE}
# Load the ToothGrowth data
data("ToothGrowth")

# Display the structure and first few rows of the data
str(ToothGrowth)
head(ToothGrowth)

```

The ToothGrowth dataset contains the following variables:

len: Tooth length
supp: Supplement type (VC or OJ)
dose: Dose in milligrams per day

## Comparing Tooth Growth by Supplement and Dose
We will compare tooth growth by supplement type and dose using t-tests and confidence intervals.

```{r, echo=FALSE}
# Perform t-test to compare tooth growth by supplement type
t_test_supp <- t.test(len ~ supp, data = ToothGrowth)

# Display the results
t_test_supp


```

## Compare Tooth Growth by Dose

```{r, echo=FALSE}
# Convert dose to a factor for comparison
ToothGrowth$dose <- as.factor(ToothGrowth$dose)

# Perform ANOVA to compare tooth growth by dose
anova_result <- aov(len ~ dose, data = ToothGrowth)

# Display the ANOVA results
summary(anova_result)

# Perform pairwise t-tests with Bonferroni correction
pairwise_t_test <- pairwise.t.test(ToothGrowth$len, ToothGrowth$dose, p.adjust.method = "bonferroni")

# Display the pairwise t-test results
pairwise_t_test

```


## Results 

# Summary of Tooth Growth by Supplement Type
The t-test results show that there is a significant difference in tooth growth between the two supplement types (VC and OJ).

# Summary of Tooth Growth by Dose

The ANOVA results indicate that there is a significant difference in tooth growth among different doses. The pairwise t-test results with Bonferroni correction show which specific doses differ significantly from each other.

## Conclusions 
From our analysis, we conclude:

There is a significant difference in tooth growth between the two supplement types (VC and OJ).
There is a significant difference in tooth growth among different doses.

# Assumptions 
The t-tests assume that the tooth length data is normally distributed within each group.
The ANOVA assumes homogeneity of variances and normality of residuals.

## Appendix 

```{r, echo=FALSE}
# Display the code used in the analysis
data("ToothGrowth")
str(ToothGrowth)
head(ToothGrowth)
summary(ToothGrowth)
t_test_supp <- t.test(len ~ supp, data = ToothGrowth)
t_test_supp
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
anova_result <- aov(len ~ dose, data = ToothGrowth)
summary(anova_result)
pairwise_t_test <- pairwise.t.test(ToothGrowth$len, ToothGrowth$dose, p.adjust.method = "bonferroni")
pairwise_t_test


```

