---
title: "Simple Prediction App Pitch"
author: "jameel"
date: "`r Sys.Date()`"
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shiny)
```



## Slide 1: Title
## Simple Prediction App

## Slide 2: Introduction

The Simple Prediction App takes a numeric input from the user and provides a prediction by multiplying the input value by 2.

## Slide 3: Features

- User input via text box
- Simple prediction algorithm
- Real-time output display

## Slide 4: How to Use

1. Enter a numeric value in the text box.
2. Click 'Predict' button.
3. View the predicted value displayed on the main panel.

## Slide 5: Demo


```{r Demo}
shinyApp(
  ui = fluidPage(
    titlePanel("Advanced Prediction App"),
    sidebarLayout(
      sidebarPanel(
        textInput("text", "Enter a value:", "50"),
        radioButtons("operation", "Choose an operation:",
                     choices = list("Multiply by 2" = "multiply", "Square the value" = "square")),
        actionButton("predict", "Predict"),
        br(),
        p("Enter a numeric value in the text box, choose an operation, and click 'Predict' to see the result."),
        p("This application can either multiply the entered value by 2 or square the value based on your choice.")
      ),
      mainPanel(
        h3("Prediction Output:"),
        verbatimTextOutput("prediction")
      )
    )
  ),
  server = function(input, output) {
    observeEvent(input$predict, {
      output$prediction <- renderText({
        input_value <- as.numeric(input$text)
        operation <- input$operation
        prediction <- if (operation == "multiply") {
          input_value * 2
        } else {
          input_value^2
        }
        paste("Predicted value:", prediction)
      })
    })
  }
)


```
