---
title: "Advanced Text Prediction Model"
author: "jameel"
date: "2024-05-28"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```

## Title Slide

- **Advanced Text Prediction Model**
- **Improving Predictive Accuracy with N-Grams and RNNs**
- **Author:** Jameel
- **Date:** 2024-05-27

## How the Model Works

- **Data Preprocessing:**
  - Tokenization: Converting text into tokens.
  - Profanity Filtering: Removing inappropriate words.

- **N-Gram Models:**
  - Unigram, Bigram, Trigram models for predicting next words.
  - Techniques: Pruning, Kneser-Ney smoothing for better performance.

- **Advanced Models:**
  - RNNs with LSTM layers to capture long-range dependencies and context.


## Predictive Performance Visualization1
```{r perf}
knitr::opts_chunk$set(echo = FALSE)

# Generate accuracy plot
model_names <- c("Unigram", "Bigram", "Trigram")
accuracy <- c(0.60, 0.70, 0.75)  # Adjust these values based on your model's performance

accuracy_df <- data.frame(Model = model_names, Accuracy = accuracy)
accuracy_plot <- ggplot(accuracy_df, aes(x = Model, y = Accuracy)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() +
  ggtitle("Model Accuracy Comparison")

# Display the plot
print(accuracy_plot)


# Generate perplexity plot
perplexity <- c(250, 150, 120)  # Adjust these values based on your model's performance
perplexity_df <- data.frame(Model = model_names, Perplexity = perplexity)
perplexity_plot <- ggplot(perplexity_df, aes(x = Model, y = Perplexity)) +
  geom_bar(stat = "identity", fill = "darkorange") +
  theme_minimal() +
  ggtitle("Model Perplexity Comparison")

# Display the plot
print(perplexity_plot)


```



## Conclusion

- **Benefits:**
  - Improved predictive accuracy.
  - Efficient computation with advanced models.
  - User-friendly app interface.

- **Call to Action:**
  - If you're interested in investing in our predictive technology.
  - Contact us: jameel@example.com